\documentclass[10pt]{scrartcl}
\usepackage{geometry}
\usepackage[parfill]{parskip}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{listings}
\lstset{
  frame=single,language=bash,
  morekeywords={od, in, foreach, let, end, and, or, proc, while},
   basicstyle=\footnotesize,
  escapechar=\@,
  basicstyle=\footnotesize, frame=tb,
  numbers=left,
  stepnumber=2,
  numbersep=5pt, 
  numberstyle=\tiny\color{mygray},
  xleftmargin=.2\textwidth, xrightmargin=.2\textwidth
}
\definecolor{bluekeywords}{rgb}{0.13,0.13,1}
\definecolor{greencomments}{rgb}{0,0.5,0}
\definecolor{redstrings}{rgb}{0.9,0,0}
\definecolor{mygray}{rgb}{0.2,0.2,0.2}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{02257 - Applied functional programming}
\subtitle{Project1 - Interpreter}
\author{Anna Maria Walach - \textit {s121540@student.dtu.dk} \\ Kim Rostgaard Christensen - \textit {s084283@student.dtu.dk}}
\begin{document}
\maketitle
\section{Object types}
There are three main categories of objects defined in the language:
\begin{itemize}
\item procedures, represented by tuple of arguments' names, procedure environment and set of procedure statements
\item arrays, represented by tuple of length and values stored in F\# array
\item primitive values: int, string, boolean
Complex structures (procedure, array) can never be send as a value to the function - always as reference. 
\end{itemize}
\subsection{Array handling}
Array variable points to reference, that store the whole array object. Values can be assigned to array cells by calling \texttt{array[index] = newValue}. Each array has attribute "length" that can be access by \texttt{array.length}.
\section{Recursive and non-recursive procedures}
We decided to not explicitly implement the recursive procedures in our interpreter. Instead, each procedure (regardless the \texttt{rec} keyword) can include recursive calls. \\
Procedure need to be always structured with let... in ... end, unless they contain only one statement (including \texttt{return}).
\subsection{Return statements}
Procedure can, but does not have to, return a value or reference. Reference to local value can be returned. Another arrays and procedures are always returned as references, not objects. 
\subsection{Parameter passing}
The language support two ways of passing the primitive (string, int, bool) parameters to function: by reference and by value. 
\begin{lstlisting}
let x : 5; y : 10;
	proc sum (a, b)
		a := +(!a,!b)
in
	call sum(x,y);
	call sum(!x,!y)
end
\end{lstlisting}
In the first case, described by \texttt{sum(x,y)}, we're sending the references to the function, so after this line, \texttt{x := 5}. After calling \texttt{sum(!x, !y)}, \texttt{x} value remains unchanged, as only the content of values is send to the procedure.
\section{Control statements}
\subsection{Conditionals}
Conditionals are implemented, but only for boolean expressions on integer types. The if else statement is structured in following way:
 \begin{lstlisting}
if (condition) then
	// do this
else //optional
	// do something else
fi
\end{lstlisting}
\subsection{Loops}
\subsubsection{While loop}
While loop is supported in a form:
 \begin{lstlisting}
while (condition) do
  // do something
od
\end{lstlisting}
\subsubsection{Foreach loop}
We've implemented a basic foreach language construct that, within the interpreter, translates into a while loop. An example f a foreach construction can be seen below.
  \begin{lstlisting}
foreach var in collection do
  // statements where var is accessible
od
\end{lstlisting}
\section{Infix logic operators}
Initially we wanted to extend the language with infix versions the implemented operators. However, to avoid breaking already implemented test cases, we instead added two new operators -- the logic AND and OR. This enables the following syntax;
\begin{lstlisting}
let
  boolvar :  <>(x,5) or <(y,10) and >(y,0)
in
  ...
end
\end{lstlisting}
Parsing and evaluation works, but unfortunately we have a shift-reduce conflict in both logic expression that we have not been able to resolve.
\section{Test cases}

\end{document}  
